|

### 2021-05-20 03-Chp-Systematic-Review.Rmd
End of day: -3892
End of day: -4243

* Smeet et al. reference in the CPRD analysis
* Larsson - 10.1136/bmj.j5375 [@larsson2017b]

@@

@@

>>>>

* Smeet et al. reference in the CPRD analysis
* Larsson - 10.1136/bmj.j5375 [@larsson2017b]
* Statins and the risk of dementia Jick

@@

Following guidance for a dose response meta-analysis, we picked m

<!-- TODO Cross reference with the data extraction issues here - i.e. not providing qratile cut-off values, not poviding numbers per group, etc. -->

@@

<!--- Comment here on issues involved in extracting data --->

### 2021-05-20 04-Chp-CPRD-Analysis.Rmd
End of day: -4243

<!----------------------------------------------------------------------------->

@@

<!----------------------------------------------------------------------------->

@@

<!----------------------------------------------------------------------------->
(ref:readExample-caption) Example of CPRD Read code hierarchy, showing how "Dementia in Alzheimer’s disease with late onset" is nested under the top-level of "Mental disorders". Broad topics are specified using the initial two bytes, while subsequent bytes are used to define specific conditions and context.

@@

```{r readExample-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

Using this system, lists of relevant codes for each of the index events, exposures and outcomes described in the sections below were created.

@@

We performed a prospective cohort study using data from the CPRD. Our initial sample included all participants registered at a participating practice between 1 January 1995 and 29 February 2016 who had a flag for "research quality" data. All events of interest were identified using predetermined code lists, which are available for inspection.

<!-- NOTE If protocol was written by Venexia, how do I reference this? -->

@@

<!--- TODO Include short bit on follow-up here, describing how and why it was defined --->

We considered five outcomes as part of this analysis: probable  Alzheimer’s disease, possible Alzheimer’s disease, vascular dementia, other dementia, and a composite all-cause dementia outcome (Supplementary Figure 1). When two or more outcomes were coded in a participant’s clinical record, a decision tree was used to differentiate between them (Supplementary Figure 1). The diagnosis date of the outcome was determined by the first record of a relevant code.

Participants were followed up, and were censored at the earliest of date of diagnosis, death, date of transfer out of the study cohort or end of follow-up ( <!--- TODO Get date here ---> )

As a positive control outcome, we included cornonary artery disease. Statins have a known effect on this outcome on the basis of randomised controlled trials. __[CITATION NEEDED]__ We also used a negative control of back pain to investigate the potential for residual confounding.

<!--- TODO need more information on how the controls were modelled - e.g. using the same covariates as in the main model or not? --->

<!--- TODO Will also need some more here on why there were chosen --->

@@

<!--- TODO Need a lot more here - see other thesis for example of how covariates are described --->

@@

All covariates were determined at index and definitions for each can be found in the archived data repository reflecting this study <!--- TODO link to Zenodo here --->

@@

<!--- TODO Need to complete paragraph below --->

<!--- Failing to adjust for missing data can. . .  --->

@@

<!--- Review materials from multiple imputation course --->

@@

<!--- TODO Finish paragraph from fragments below --->

@@

<!--- TODO Need to actually run these models, but the important things is that they can be run on the unimputed data! --->

&nbsp;
<!----------------------------------------------------------------------------->

@@

<!----------------------------------------------------------------------------->
(ref:cprd-dag-1-cap) __Directed acyclic graph showing assumed relationship between exposure and covariate:__

(ref:cprd-dag-1-scap) Directed acyclic graph showing assumed relationship between exposure and covariate

```{r cprd-dag-1, echo = FALSE, results="asis", fig.cap='(ref:cprd-dag-1-cap)', out.width='100%', fig.scap='(ref:cprd-dag-1-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prisma_flow.png"))
```
<!----------------------------------------------------------------------------->

@@

<!--- TODO Discuss problems introduced by censoring on death --->

<!----------------------------------------------------------------------------->
(ref:immortalTimeBias-cap) __shortcap:__

@@

```{r immortalTimeBias, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:immortalTimeBias-cap)', out.width='100%', fig.scap='(ref:immortalTimeBias-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prisma_flow.png"))
```
<!----------------------------------------------------------------------------->

<!-- ### Sensitivity analyses -->

<!--- Described sensitivity analyses reported below here --->

<!--- Make sure to include age as a sensitivity analysis, given that that is the purpose of the IPD analysis --->

@@

Of the `r comma(attrition[1,1])` participants included in our extract, `r comma(attrition[14,1])` met the inclusion criteria (Figure \@ref(fig:cprdFlowchart)), with a total follow-up of `r comma(total_followup)` <!--- TODO make this reactive! ---> patient years at risk.

@@

<!----------------------------------------------------------------------------->
(ref:cprdFlowchart-cap) __Attrition of CPRD participants as the eligibility criteria were applied:__

@@

<!--- This figure needs to have comma seperated numbers in order to agree with text. --->

```{r cprdFlowchart, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdFlowchart-cap)', out.width='100%', fig.scap='(ref:cprdFlowchart-scap)'}
knitr::include_graphics(file.path("figures","cprd-analysis","cohort_attrition.png"))
```
<!----------------------------------------------------------------------------->

@@

The median age at index was `r age_text` and participants were followed up for a median of `r fu_text`. During follow-up, an all-cause dementia diagnosis was recorded for `r comma(p1[1,2])` patients  (`r comma(p1[2,2])` probable AD, `r comma(p1[3,2])` possible AD, `r comma(p1[4,2])` vascular dementia, `r comma(p1[5,2])` other dementia; <!--- see Supplementary Figure 1 --->). The distribution of baseline characteristics across the drug classes can be seen in Table \@ref(tab:cprdCharacteristics-table).

\blandscape
<!----------------------------------------------------------------------------->
(ref:cprdCharacteristics-caption) Patient characteristics by drug class

@@

```{r cprdCharacteristics-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape

<!--- Table 1 is missing type 2 diabetes --->

@@

<!--- TODO Describe impact of SSA table - is the fact that many people stopped statin treatment (~10%) an issue? --->

@@

<!----------------------------------------------------------------------------->
(ref:cprdSSA-caption) Participants who stopped, switched or added treatements by initial treatment type

@@

```{r cprdSSA-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

<!----------------------------------------------------------------------------->
(ref:cprdPrimary-cap) __Results from primary analyses of CPRD data:__

@@

```{r cprdPrimary, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdPrimary-cap)', out.width='100%', fig.scap='(ref:cprdPrimary-scap)'}
knitr::include_graphics(file.path("figures","cprd-analysis","fp_p1.jpeg"))
```
<!----------------------------------------------------------------------------->

@@

For the composite all-cause dementia outcome, we found treatment with a lipid-regulating agent was associated with a slightly increased risk (`r anydem_text`), but the magnitude of the association was not as extreme as that observed for the vascular dementia subgroup. There was also some evidence that fibrates were associated with increased risk of all-cause dementia (`r anydem_fib_text`).

@@

When stratifying based on year of entry to the cohort, we observed no variation in risk by time period in any subgroup except for probable Alzheimer’s disease (Figure \@ref(hold)). Removing participants coded as female aged 55 and under at index from our analysis had minimal effect on our estimates (Figure \@ref(hold)).

<!--- TODO Actually run this analysis --->

<!--- TODO Sort out the mess that is below --->

@@

<!--- I actually think it would be better to include these results not in the appendix --->

<!--- By multiple imputation versus complete case analysis --->

@@

<!--- By time varying confounder --->

@@

<!-- This means we can find the most recent value of the covariate before the switch and apply a marginal structural model approach, filling all values for that variable before the most recent measure with the baseline measurement, and all after the most recent meausure with the value of the most recent measure (on the basis that you won't go from having CVD back to not having CVD). -->

@@

<!-- ### Controls  -->

<!--- TODO I think this is maybe dealth with at the end of the "Exposures" section, but better to have it here? --->

<!-- Backpain -->

<!-- Coronary artery disease -->

<!----------------------------------------------------------------------------->
&nbsp;

@@

<!--- To what extent should I cross-reference with the other literature here? --->

<!--- Maybe it would be better to signpost towards the triangulation chapter where the different evidence streams are combined --->

@@

<!--- TODO Why are fibrates different? --->

<!--- TODO Be explicit here about the possibility of publication bias in the vascular dementia results --->

@@

Several previous studies have demonstrated that statins are the overwhelmingly favoured methods of cholesterol control. <!--- TODO citation needed --->

@@

To our knowledge, there is no previous study of the effect of preventative treatment with ezetimibe on any dementia outcome, and so we cannot compare our unexpected finding that treatment with the drug associated with an increased risk of the vascular and other dementia outcomes.

<!----------------------------------------------------------------------------->
&nbsp;

@@

<!--- TODO I actually don't think Strengths and Limitations are the best way to describe this! --->

<!-- TODO A major strength of this analysis is the size of the included cohort and the length of follow-up that the use of electronic health records allowed. In addition, we followed users and non-users from a common index date, using a time-updating treatment indicator to correctly assign time-at-risk to the exposed and unexposed groups. -->

@@

This analysis may be subject to confounding by indication, which occurs when factors that affect whether a participant is exposed also affect their outcome. We attempted to address this by limiting inclusion to those either prescribed or “at risk” of being prescribed, which was determined using an elevated test result. We also adjusted for several additional potential confounding variables. However, the negative control analysis of back pain demonstrated a harmful association with lipid-regulating agent use, indicting that this findings may be biased by residual confounding. Important confounding variables for which we have not adjusted could include genetic factors. A recent preprint of a study in the UK Biobank demonstrated that an Alzheimer’s disease polygenic risk score was associated with an increased risk of unspecified Alzheimer’s and vascular dementia, and also with an increased frequency of self-reported raised cholesterol levels, a diagnosis of hypercholesterolaemia, and a history of taking lipid-regulating agents such as statins or ezetimibe.[@korologou-linden2020] This finding, combined with the potential for differential misclassification between Alzheimer’s disease and vascular dementia, could explain part of the observed association between lipid-regulating agents and vascular dementia.

<!--- TODO A further limitation is the possibility of time-varying confounding  Either deal with this issue here, or keep the section above dealing with time-varying confounding. --->

@@

<!----------------------------------------------------------------------------->
&nbsp;

@@

<!--- TODO I will definitely have to mention this earlier in the chapter, as it is bad for to try and spring methods on people --->

@@

<!--- TODO check that this is actually the case, that when you use the codelist from Smeeth, you get a different case-mix. Present it as a table. --->
<!--- Will need to present the results for this! --->
<!-- This, in addition to unexplained statistical issues such as adjusting for post-baseline covariates, means I am sceptical of the value of comparing our findings with Smeeth. -->
<!-- IN addition, this analysis used propensity s. IN addition, an analytical approach which appeared to adjust for key variables likely to be on the analytical pathway between   -->

@@

<!----------------------------------------------------------------------------->
(ref:smeethComparison-caption) Comparison of code lists used in this analysis and in Smeeth et al.

@@

```{r smeethComparison-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

<!--- This hasn't been done! Plus I do not think this is the case --->

<!----------------------------------------------------------------------------->

<!----------------------------------------------------------------------------->
<!-- &nbsp; -->

@@

<!----------------------------------------------------------------------------->
&nbsp;

### Enabling easy synthesis of this analysis

In light of my own experiences in attempting to extract information for papers assessing preventative measures, as documented in Section \@ref(sys-rev-open-data), the outputs from this analysis are described in detail in both. IN line with this, this analyses has been reported in line with the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) reporting guidelines[@vandenbroucke2007] (see Appendix \@ref(hold) for the STROBE checklist). <!--- Maybe don't include the whole thing, or apply it to this thesis rather than the paper --->

@@

All code and summary statistics (i.e. the tables presented in this chapter, plus summary tables of effect estimates) are readily available in machine readable formats (i.e. as comma separated values, or CSV, files) from the Github repository for this project. <!--- TODO Will need to cite this, and may need to archive on Zenodo. --->

<!----------------------------------------------------------------------------->
&nbsp;

@@

 This chapter attempted to account for important sources of bias and provide a comparison with other available literature, as identified in the systematic review presented in Chapter \@ref(sys-rev-heading). However, there is a strong potential for unmeasured confounding, misclassification and reverse causation, which raises questions about our findings, in particular the unexpected increase in risk of vascular dementia associated with statin use. <!-- TODO Finish This is supported by our findings for the negative and positive controls used. -->

@@

<!--- TODO In my most recent discussion with supervisors, I discussed that a long term RCT isn't encouraged by the results --->

<!--- TODO Need better linkage with next chapter and triangulation --->

>>>>

<!--------------------------------------------------------------------------->

@@

<!-- TODO Add amendments to Appendix -->

<!--------------------------------------------------------------------------->

@@

<!--------------------------------------------------------------------------->
(ref:readExample-caption) Example of CPRD Read code hierarchy, showing how "Dementia in Alzheimer’s disease with late onset" is nested under the top-level of "Mental disorders". Broad topics are specified using the initial two alpha-numeric characters of the Read code, while subsequent characters are used to define specific conditions and context.

@@

```{r readExample-table, message=FALSE, results="asis", echo = FALSE}
```
<!--------------------------------------------------------------------------->

@@

Using this system, lists of relevant codes for each of the index events, exposures and outcomes used in the analysis were created.

@@

This analysis was a prospective cohort study using data from the CPRD. It included all participants registered at a participating practice between 1 January 1995 and 29 February 2016 who had a flag for "research quality" data. Records pre-dating the 1995 cutoff were included in the original CPRD extract but were excluded from the analysis as records are expected to be more complete and reliable after this date.[@wolf2019] Additionally, individuals with less than 12 months of continuous records prior to cohort entry were excluded, making the effective start date of the cohort the 1st January 1996.

All events of interest were identified using predetermined code lists, which are available for inspection from the archived repository accompanying this analysis (data/code availability is discussed in Section \@ref(cprd-data-avail)).

<!-- QUESTION If protocol was written by Venexia, how do I reference this? -->

@@

We considered five outcomes as part of this analysis: probable  Alzheimer’s disease, possible Alzheimer’s disease, vascular dementia, other dementia, and a composite all-cause dementia outcome (Supplementary Figure 1). When two or more outcomes were coded in a participant’s clinical record, a decision tree was used to differentiate between them (see Figure ). The diagnosis date of the outcome was determined by the first record of a relevant code.

Participants were followed up, and were censored at the earliest of date of diagnosis, death, date of transfer out of the study cohort or end of follow-up (29th February 2016).

As a positive outcome control, we included ishemic heart disease. Statins have a known effect on this outcome on the basis of randomised controlled trials. __[CITATION NEEDED]__ We also used a negative outcome control of back pain to investigate the potential for residual confounding.

<!-- TODO need more information on how the controls were modelled - e.g. using the same covariates as in the main model or not? I.e. should those with baseline cardiovascular conditions be excluded from the IHD/MI analysis? -->

<!-- TODO Will also need some more here on why there were chosen -->

The inclusion of control outcomes are a useful technique to assess the strength of uncontrolled confounding in an analysis.[@lipsitch2010] Despite observational analyses suggesting a link between statins and muscle pain, systematic reviews of the adverse events of statin use[@collins2016a] and N-of-1 trials explicitly exploring the effect of statin use on muscle pain[@herrett2021] have found no evidence of an effect. As such, muscular backpain was used as a negative control outcome in this analysis.

<!----------------------------------------------------------------------------->
(ref:negativeOutcome-cap) Causal diagram showing relationship between exposure [X], outcome [Y], confounders (measured [C] and unmeasured [U]) and an ideal negative outcome [N]:

(ref:negativeOutcome-scap) DAG of ideal negative outcome

```{r negativeOutcome, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:negativeOutcome-cap)', out.width='100%', fig.scap='(ref:negativeOutcome-scap)', message=FALSE, warning = FALSE}
library(dagitty)
library(ggdag)

ggplot2::ggsave(filename = file.path("figures","cprd-analysis","negativeOutcome.png"),
  dagitty::dagitty("dag {
    U -> N
    C -> N
    X <- U -> Y
    X <- C -> Y
    X -> Y
    bb =\"-2.368,-2.972,2.597,2.726\"
    C [pos=\"-0.3,0\"]
    N [pos=\"0.5,1\"]
    U [pos=\"-0.15,1.5\"]
    X [exposure,pos=\"0,-1\"]
    Y [outcome,pos=\"0.5,-1\"]
  }"
) %>% ggdag_classic() +
  theme_dag()

,
height = 3
)

knitr::include_graphics(file.path("figures","cprd-analysis","negativeOutcome.png"))

```
<!----------------------------------------------------------------------------->

@@

<!-- TODO Need a lot more here - see other thesis for example of how covariates are described -->

@@

All covariates were determined at index and definitions for each can be found in the archived data repository reflecting this study <!-- TODO link to Zenodo here -->

@@

<!-- TODO Need to complete paragraph below -->

<!-- Failing to adjust for missing data can. . .  -->

@@

<!-- Review materials from multiple imputation course -->

@@

<!-- TODO Finish paragraph from fragments below -->

@@

<!-- TODO Need to actually run these models, but the important things is that they can be run on the unimputed data! -->

&nbsp;
<!--------------------------------------------------------------------------->

@@

@@

<!-- TODO Discuss problems introduced by censoring on death -->

<!--------------------------------------------------------------------------->
(ref:immortalTimeBias-cap) __shortcap:__

@@

```{r immortalTimeBias, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:immortalTimeBias-cap)', out.width='100%', fig.scap='(ref:immortalTimeBias-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prisma_flow.png"))
```
<!--------------------------------------------------------------------------->  
### Sensitivity analyses

<!-- TODO Described sensitivity analyses reported below here -->

In addition to the primary analysis

Statins are contra-indicated in pregnant patients, and so I preformed a sensitivity analysis in patients aged 55 and coded as female.

We also prod

<!-- Make sure to include sex as a sensitivity analysis, given that that is the purpose of the IPD analysis -->

@@

```{r attritionFigure, include = FALSE}
```

Of the `r comma(attrition[1,1])` participants included in our extract, `r comma(attrition[14,1])` met the inclusion criteria (Figure \@ref(fig:cprdFlowchart)), with a total follow-up of `r comma(total_followup)` patient years at risk.

@@

<!--------------------------------------------------------------------------->
(ref:cprdFlowchart-cap) __Attrition of CPRD participants as the eligibility criteria were applied:__

@@

<!-- This figure needs to have comma seperated numbers in order to agree with text. -->

```{r cprdFlowchart, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdFlowchart-cap)', out.width='100%', fig.scap='(ref:cprdFlowchart-scap)'}
knitr::include_graphics(file.path("figures","cprd-analysis","cohort_attrition_test.png"))
```
<!--------------------------------------------------------------------------->

@@

The median age at index was `r age_text` and participants were followed up for a median of `r fu_text`. During follow-up, an all-cause dementia diagnosis was recorded for `r comma(p1[1,2])` patients  (`r comma(p1[2,2])` probable AD, `r comma(p1[3,2])` possible AD, `r comma(p1[4,2])` vascular dementia, `r comma(p1[5,2])` other dementia; <!-- see Supplementary Figure 1 -->). The distribution of baseline characteristics across the drug classes can be seen in Table \@ref(tab:cprdCharacteristics-table).

\blandscape
<!--------------------------------------------------------------------------->
(ref:cprdCharacteristics-caption) Patient characteristics by drug class

@@

```{r cprdCharacteristics-table, message=FALSE, results="asis", echo = FALSE}
```
<!--------------------------------------------------------------------------->
\elandscape

<!-- Table 1 is missing type 2 diabetes -->

@@

<!-- TODO Describe impact of SSA table - is the fact that many people stopped statin treatment (~10%) an issue? Comment specifically in the context of a -->

<!-- TODO Described the average length of time between patients being diagnosed with  -->

@@

<!--------------------------------------------------------------------------->
(ref:cprdSSA-caption) Participants who stopped, switched or added treatements by initial treatment type

@@

```{r cprdSSA-table, message=FALSE, results="asis", echo = FALSE}
```
<!--------------------------------------------------------------------------->

@@

<!--------------------------------------------------------------------------->
(ref:cprdPrimary-cap) __Results from primary analyses of CPRD data:__

@@

```{r cprdPrimary, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdPrimary-cap)', out.width='100%', fig.scap='(ref:cprdPrimary-scap)'}
knitr::include_graphics(file.path("figures","cprd-analysis","fp_p1.jpeg"))
```
<!--------------------------------------------------------------------------->

@@

For the composite all-cause dementia outcome, I found treatment with a lipid-regulating agent was associated with a slightly increased risk (`r anydem_text`), but the magnitude of the association was not as extreme as that observed for the vascular dementia subgroup. There was also some evidence that fibrates were associated with increased risk of all-cause dementia (`r anydem_fib_text`).

@@

When stratifying based on year of entry to the cohort, I observed no variation in risk by time period in any subgroup except for probable Alzheimer’s disease (Figure \@ref(hold)). Removing participants coded as female and aged 55 and under at index from our analysis had minimal effect on our estimates (Figure \@ref(hold)).

<!-- TODO Actually run this analysis -->

<!-- TODO Sort out the mess that is below -->

@@

<!-- I actually think it would be better to include these results not in the appendix -->

<!-- By multiple imputation versus complete case analysis -->

@@

<!-- By time varying confounder -->

@@

<!-- This means I can find the most recent value of the covariate before the switch and apply a marginal structural model approach, filling all values for that variable before the most recent measure with the baseline measurement, and all after the most recent meausure with the value of the most recent measure (on the basis that you won't go from having CVD back to not having CVD). -->

@@

<!--------------------------------------------------------------------------->
&nbsp;

@@

<!-- To what extent should I cross-reference with the other literature here? -->

<!-- Maybe it would be better to signpost towards the triangulation chapter where the different evidence streams are combined -->

@@

<!-- TODO Why are fibrates different? -->

<!-- TODO Be explicit here about the possibility of publication bias in the vascular dementia results -->

<!-- QUESTION How much should I be cross-referencing with the systematic review at this stage? Or should I keep it self-contained -->

<!-- TODO Discuss impact of sensitivity analysis results, particularly -->

The effect estimate seems to increase as the casemix ratio of Alzheimers:vascular dementia decreases.

The fact that a similar increased risk of IHD is observed in those taking statins likely means that those who end up taking statins (or another lipid regulating agent) are inherently different from those who do not in some way that this analysis does not account for.

Likely confounding by indication

Confounding by indication occurs when an unobserved variable both prompts exposure to the intervention and modifies a participants risk of the outcome. In this case, this confounder would prompt prescription of statins (or another lipid regultating agent) but also represent a vascular risk factor that contributes to the development of the control outcome (ischemic heart disease) and also dementia outcomes with an increasing casemix of vascular origin (e.g. the vascular dementia and other dementia endpoints)

In causal inference language, statins and dementia are said to be d-connected (see Figure \@ref(fig:indicationBias)), as there is an open "backdoor" path between them, via the

Alternatively, rather than a single variable, it is likely that those that take statins are different in a myriad of ways, each of which contributes a

A common example of this bias

While this analysis has attempted to adjust for several baseline variables that my prompt

Possible further analysis techniques would be the use of

<!----------------------------------------------------------------------------->
(ref:indicationBias-cap) __Confounding by indication directed acyclic graph:__

(ref:indicationBias-scap) Confounding by indication directed acyclic graph

```{r indicationBias, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:indicationBias-cap)', out.width='100%', fig.scap='(ref:indicationBias-scap)', message = FALSE}

ggplot2::ggsave(filename = file.path("figures","cprd-analysis","indicationBias.png"),
dagitty::dagitty("dag {
  Statin [exposure,pos=\"0.000,0.000\"]
  \"Vascular outcomes\" [outcome,pos=\"2.000,0.000\"]
  \"Uncontrolled variable\" [pos=\"1.000,1.000\"]
  \"Uncontrolled variable\" -> Statin
  \"Uncontrolled variable\" -> \"Vascular outcomes\"
}") %>% ggdag_classic(size = 6) +
  xlim(-0.5,3) +
  geom_text(data = data.frame(x = 0.316392393234329,
                              y = 0.537705383514239,
                              label = "Prompts"),
            mapping = aes(x = x,
                          y = y,
                          label = label),
            angle = 0L,
            lineheight = 1L,
            hjust = 0.5,
            vjust = 0.5,
            colour = "red",
            family = "sans",
            fontface = "bold",
            inherit.aes = FALSE,
            show.legend = FALSE) +
  geom_text(data = data.frame(x = 1.65891420807966,
                              y = 0.537705383514239,
                              label = "Causes"),
            mapping = aes(x = x,
                          y = y,
                          label = label),
            angle = 0L,
            lineheight = 1L,
            hjust = 0.5,
            vjust = 0.5,
            colour = "red",
            family = "sans",
            fontface = "bold",
            inherit.aes = FALSE,
            show.legend = FALSE) +
  geom_curve(data = data.frame(x = 0.220612676359188,
                               y = -0.000788829298738025,
                               xend = 1.45,
                               yend = -0.00157828084913909),
             mapping = aes(x = x,
                           y = y,
                           xend = xend,
                           yend = yend),
             linetype = "dashed",
             angle = 0L,
             size = 0.6,
             colour = "black",
             curvature = 0,
             arrow = structure(list(angle = 0,
                                    length = structure(0,
                                                       class = "unit",
                                                       valid.unit = 2L,
                                                       unit = "inches"),
                                    ends = 2L,
                                    type = 2L),
                               class = "arrow"),
             inherit.aes = FALSE,
             show.legend = FALSE) +

  theme_dag(),
width = 8,
height = 3)

knitr::include_graphics(file.path("figures","cprd-analysis","indicationBias.png"))
```
<!----------------------------------------------------------------------------->

Similar suprising findings for the effect of statins on

In essence, this

@@

Several previous studies have demonstrated that statins are the overwhelmingly favoured methods of cholesterol control. <!-- TODO citation needed -->

@@

To our knowledge, there is no previous study of the effect of preventative treatment with ezetimibe on any dementia outcome, and so I cannot compare our unexpected finding that treatment with the drug associated with an increased risk of the vascular and other dementia outcomes.

<!--------------------------------------------------------------------------->
&nbsp;

@@

<!-- TODO I actually don't think Strengths and Limitations are the best way to describe this! -->

<!-- TODO A major strength of this analysis is the size of the included cohort and the length of follow-up that the use of electronic health records allowed. In addition, I followed users and non-users from a common index date, using a time-updating treatment indicator to correctly assign time-at-risk to the exposed and unexposed groups. -->

@@

This analysis may be subject to confounding by indication, which occurs when factors that affect whether a participant is exposed also affect their outcome. We attempted to address this by limiting inclusion to those either prescribed or “at risk” of being prescribed, which was determined using an elevated test result. We also adjusted for several additional potential confounding variables. However, the negative control analysis of back pain demonstrated a harmful association with lipid-regulating agent use, indicting that this findings may be biased by residual confounding. Important confounding variables for which I have not adjusted could include genetic factors. A recent preprint of a study in the UK Biobank demonstrated that an Alzheimer’s disease polygenic risk score was associated with an increased risk of unspecified Alzheimer’s and vascular dementia, and also with an increased frequency of self-reported raised cholesterol levels, a diagnosis of hypercholesterolaemia, and a history of taking lipid-regulating agents such as statins or ezetimibe.[@korologou-linden2020] This finding, combined with the potential for differential misclassification between Alzheimer’s disease and vascular dementia, could explain part of the observed association between lipid-regulating agents and vascular dementia.

<!-- TODO A further limitation is the possibility of time-varying confounding  Either deal with this issue here, or keep the section above dealing with time-varying confounding. -->

@@

<!--------------------------------------------------------------------------->
&nbsp;

@@

<!-- TODO I will definitely have to mention this earlier in the chapter, as it is bad for to try and spring methods on people -->

@@

<!-- TODO check that this is actually the case, that when you use the codelist from Smeeth, you get a different case-mix. Present it as a table. -->
<!-- Will need to present the results for this! -->
<!-- This, in addition to unexplained statistical issues such as adjusting for post-baseline covariates, means I am sceptical of the value of comparing our findings with Smeeth. -->
<!-- IN addition, this analysis used propensity s. IN addition, an analytical approach which appeared to adjust for key variables likely to be on the analytical pathway between   -->

@@

<!--------------------------------------------------------------------------->
(ref:smeethComparison-caption) Comparison of code lists used in this analysis and in Smeeth et al.

@@

```{r smeethComparison-table, message=FALSE, results="asis", echo = FALSE}
```
<!--------------------------------------------------------------------------->

@@

<!-- This hasn't been done! Plus I do not think this is the case -->

<!--------------------------------------------------------------------------->

<!--------------------------------------------------------------------------->
<!-- &nbsp; -->

@@

<!--------------------------------------------------------------------------->
&nbsp;

### Enabling easy synthesis of this analysis {#cprd-data-avail}

In light of my own experiences in attempting to extract information for papers assessing preventative measures, as documented in Section \@ref(sys-rev-open-data), the outputs from this analysis are described in detail in both.

All code, Read codelist and summary statistics (i.e. the tables presented in this chapter, plus summary tables of effect estimates) are readily available in machine readable formats (i.e. as comma separated values, or CSV, files) from the archived repository for this project. <!-- TODO Will need to cite this, and may need to archive on Zenodo. -->

The raw data supporting this analysis is not available, as access to the CPRD data is controlled by a data monitoring committee. Interested researchers can apply directly to the  In this scenario, sharing the code represents a <!-- TODO CITATION NEEDED: Goldarce BMJ opinion -->

As such, this analyses has been reported in line with the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) reporting guidelines[@vandenbroucke2007] (see Appendix \@ref(hold) for the STROBE checklist). <!-- Maybe don't include the whole thing, or apply it to this thesis rather than the paper -->

@@

<!--------------------------------------------------------------------------->
&nbsp;

@@

 This chapter attempted to account for important sources of bias and provide a comparison with other available literature, as identified in the systematic review presented in Chapter \@ref(sys-rev-heading). However, there is a strong potential for unmeasured confounding, misclassification and reverse causation, which raises questions about our findings, in particular the unexpected increase in risk of vascular dementia associated with statin use. <!-- TODO Finish This is supported by our findings for the negative and positive controls used. A slight increas-->

@@

<!-- TODO In my most recent discussion with supervisors, I discussed that a long term RCT isn't encouraged by the results -->

<!-- TODO Need better linkage with next chapter and triangulation -->
